\phantomsection\addcontentsline{toc}{section}{Biomimetics}
\ECchapter{19}{Biomimetics}{How can engineering learn from living systems?}{ECPurple}

\ECObjectives{
\begin{itemize}
\item Distinguish biological inspiration, abstraction and direct imitation.
\item Transfer a biological working principle into a manufacturable engineering concept.
\item Compare biomimetic and conventional solutions using performance and life-cycle evidence.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Biomimetic research is influencing adhesives, lightweight structures, ventilation systems and robot
locomotion. Nature provides useful strategies, but engineering success requires abstraction,
modelling and testing rather than simply copying the appearance of an organism.
\end{engineeringnews}

\begin{ecarticle}
Biomimetics examines biological functions and transfers useful principles to engineered systems. The
process usually begins with a technical problem such as reducing drag, gripping a surface or
building a lightweight structure. Engineers then search for organisms that solve a comparable
functional problem under similar constraints.

The biological solution must be understood at the relevant scale. A bird wing is not only a curved
surface; its performance depends on flexible feathers, muscles, sensing and active control. Directly
copying its shape may therefore fail. Engineers abstract the underlying principle, create a model and
adapt it to available materials and manufacturing methods.

Surface engineering provides many examples. Gecko feet adhere through millions of microscopic hairs
that create weak intermolecular forces over a large contact area. Artificial dry adhesives reproduce
the hierarchical structure without using glue. Shark skin has rib-like textures that influence the
flow close to the surface. Similar riblets can reduce friction drag in selected conditions, although
their benefit depends on size, orientation, Reynolds number and cleanliness.

Nature also inspires structures. Bone combines a dense outer layer with a porous interior and
achieves a high stiffness-to-mass ratio. Additive manufacturing can produce lattice structures that
use the same principle. Honeycomb panels and branching networks similarly offer efficient ways to
distribute loads or fluids.

A biomimetic design is not automatically sustainable. Manufacturing a complex microstructure may
consume large amounts of energy or use materials that are difficult to recycle. Engineers must compare
the complete life cycle, durability and maintenance requirements with a conventional solution.
Biomimetics is valuable only when the biological analogy leads to a measurable improvement.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Engineering Insight,colback=yellow!10,colframe=ECOrange]
The hook-and-loop fastener was inspired by burrs attached to animal fur. Its success came from
abstracting a repeatable attachment principle that could be manufactured economically, not from
reproducing the complete plant structure.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{From biological observation to engineering validation}
\begin{center}
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=3.3mm]
\node[draw=ECPurple,fill=ECPurple!10,rounded corners,align=center] (p) {define the\\technical function};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,align=center,right=of p] (b) {observe a\\biological strategy};
\node[draw=ECBlue,fill=ECLightBlue,rounded corners,align=center,right=of b] (a) {abstract the\\working principle};
\node[draw=ECOrange,fill=ECOrange!10,rounded corners,align=center,below=of a] (m) {model and\\prototype};
\node[draw=ECRed,fill=ECRed!8,rounded corners,align=center,below=of b] (v) {test performance\\and life cycle};
\draw[->,thick] (p)--(b); \draw[->,thick] (b)--(a); \draw[->,thick] (a)--(m);
\draw[->,thick] (m)--(v); \draw[->,thick,dashed] (v)--(p);
\end{tikzpicture}%
}
\end{center}
Iteration is essential because biological materials, scales and manufacturing constraints are different.
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
Primary objective & functional transfer\\
Common design scale & microstructure\\
Typical benefit & low mass or drag\\
Necessary step & abstraction\\
Final proof & measured validation
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Illustrative drag comparison,colback=white,colframe=ECPurple]
\begin{center}
\begin{tikzpicture}
\begin{axis}[xbar,width=.94\linewidth,height=49mm,xlabel={Relative drag index},xmin=0,xmax=110,
symbolic y coords={Optimised riblets,Shark-inspired coating,Riblet texture,Smooth reference},
ytick=data,grid=major,ticklabel style={font=\scriptsize},nodes near coords]
\addplot table[x=drag,y=surface,col sep=comma]{data/EC19/drag.csv};
\end{axis}
\end{tikzpicture}
\end{center}
\footnotesize Illustrative values showing that geometry must be optimised for the operating flow and
validated against a smooth reference under identical conditions.
\end{tcolorbox}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
biomimetics & biomimétisme\\
living system & système vivant\\
working principle & principe de fonctionnement\\
hierarchical structure & structure hiérarchique\\
riblet & micro-rainure\\
adhesion & adhérence\\
porous & poreux\\
lattice structure & structure en treillis\\
stiffness-to-mass ratio & rapport rigidité/masse\\
life cycle & cycle de vie
\end{tabularx}
\end{vocabulary}

\begin{applicationexercise}
A smooth reference surface has a drag index of 100. A biomimetic riblet surface has an index of 82,
but its manufacture increases the component mass from 12.0 kg to 12.6 kg and adds 8 impact units.
\begin{enumerate}
\item Calculate the percentage reduction in drag index.
\item Calculate the percentage increase in mass.
\item Estimate the net score $S=\Delta D-2\Delta m$, where reductions are positive percentages.
\item Explain one limitation of this simplified score.
\item Define two controlled experiments and one durability test before approval.
\end{enumerate}
\end{applicationexercise}

\begin{questions}
\item Why is direct imitation often insufficient?
\item Explain how gecko feet inspire dry adhesives.
\item What engineering benefit can be obtained from bone-like structures?
\item Why must operating scale be considered for riblets?
\item How can a biomimetic product have poor environmental performance?
\item What should a fair reference solution include during testing?
\end{questions}

\begin{engineeringchallenge}
Choose a biological strategy that could improve a small wind turbine, robot gripper or protective
helmet. Define measurable requirements, abstract the biological principle and propose a manufacturable
concept. Compare it with a conventional design using a decision matrix, then specify a prototype,
three experiments, acceptance criteria and a life-cycle check.
\end{engineeringchallenge}

\begin{speaking}
Is nature always the best engineer? Discuss optimisation, evolution, materials, manufacturing and the
need for measurable evidence. Use one successful and one questionable analogy.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{AskNature -- Biomimicry examples}{https://asknature.org/}
\ECsource{Biomimicry Institute}{https://biomimicry.org/}