\phantomsection\addcontentsline{toc}{section}{Sustainable Buildings}
\ECchapter{29}{Sustainable Buildings}{Balancing comfort, energy and lifecycle carbon}{ECOrange}

\ECObjectives{
\begin{itemize}
\item Distinguish operational energy, embodied carbon and whole-life performance.
\item Apply heat-loss and heat-pump calculations to a simple building envelope.
\item Define commissioning, monitoring and acceptance criteria for a low-energy building.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Buildings create environmental impacts during material production, construction, operation, maintenance and end
of life. Sustainable design combines passive architecture, efficient systems, low-carbon materials, renewable
energy and measured verification rather than relying only on simulation labels.
\end{engineeringnews}

\begin{ecarticle}
A sustainable building must provide comfort, safety, indoor-air quality and useful space while reducing impacts
over its complete lifecycle. Operational energy is used for heating, cooling, ventilation, lighting and equipment.
Embodied carbon is associated with raw materials, manufacturing, transport, construction, replacement and demolition.
As envelopes improve and electricity becomes less carbon-intensive, embodied impacts can represent a larger share
of total emissions.

The first design principle is to reduce demand. Transmission through one envelope element can be estimated from
\[
\dot Q=UA\Delta T,
\]
where $U$ is thermal transmittance, $A$ is area and $\Delta T$ is the indoor--outdoor temperature difference.
Lower $U$-values, fewer thermal bridges and improved airtightness reduce winter losses. Orientation, glazing ratio,
external shading and thermal mass influence solar gains and summer overheating.

Airtightness must be combined with controlled ventilation. Without sufficient air renewal, humidity and pollutant
concentrations can rise. Demand-controlled ventilation may use occupancy and carbon-dioxide measurements, but
CO$_2$ is mainly an indicator of occupancy-related ventilation and not a complete measure of indoor-air quality.
Particles, volatile organic compounds, moisture and outdoor pollution may require additional sensing or filtration.

Heating and cooling equipment must be selected after the demand has been reduced. A heat pump delivers heat using
electrical work, with
\[
\mathrm{COP}=\frac{Q_{\mathrm{heat}}}{W_{\mathrm{electric}}}.
\]
COP decreases as the temperature lift increases. Low-temperature emitters, good source conditions and a well-insulated
envelope therefore improve system performance. Seasonal performance is more relevant than a single laboratory COP.

Material decisions require lifecycle assessment. Concrete offers durability and thermal mass but cement production
is carbon intensive. Timber can reduce structural mass and store biogenic carbon, yet sourcing, fire, moisture and
end-of-life scenarios must be assessed. Reusing structural components can avoid both demolition waste and new
manufacturing.

A digital model remains a prediction. The performance gap between calculated and measured consumption can result
from construction defects, incorrect controls, unexpected occupancy, weather or user behaviour. Commissioning,
sub-metering, trend analysis and corrective action are therefore part of the engineering process. A building is
successful only if it continues to satisfy requirements in operation.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Engineering insight,colback=yellow!10,colframe=ECOrange]
The lowest design consumption is not useful if controls are poorly commissioned. Requirements must therefore include
measurable operational indicators, not only simulated annual energy.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{Building performance loop}
\begin{center}
\resizebox{0.96\linewidth}{!}{%
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=3.5mm]
\node[draw=ECOrange,fill=ECOrange!10,rounded corners] (design) {design model};
\node[draw=ECBlue,fill=ECLightBlue,rounded corners,right=of design] (build) {construction};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,right=of build] (operate) {operation};
\node[draw=ECPurple,fill=ECPurple!10,rounded corners,below=of build] (meter) {metering and diagnosis};
\draw[->,thick] (design)--(build); \draw[->,thick] (build)--(operate);
\draw[->,thick] (operate)--(meter); \draw[->,thick] (meter)--(design);
\end{tikzpicture}}
\end{center}
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
First priority & reduce demand\\
Envelope equation & $\dot Q=UA\Delta T$\\
Heat-pump metric & seasonal COP\\
Air-quality proxy & CO$_2$\\
Operational proof & commissioning and metering
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Illustrative annual energy demand,colback=white,colframe=ECOrange]
\begin{center}
\begin{tikzpicture}
\begin{axis}[ybar stacked,bar width=10pt,width=0.96\linewidth,height=46mm,ylabel={Energy (kWh/m$^2$/year)},
symbolic x coords={Conventional,Efficient,Passive},xtick=data,
legend style={font=\tiny,at={(0.5,-0.24)},anchor=north,legend columns=3},
ticklabel style={font=\scriptsize},ylabel style={font=\scriptsize}]
\addplot table[x=design,y=heating,col sep=comma]{data/EC29/energy.csv};
\addplot table[x=design,y=cooling,col sep=comma]{data/EC29/energy.csv};
\addplot table[x=design,y=lighting,col sep=comma]{data/EC29/energy.csv};
\legend{Heating,Cooling,Lighting}
\end{axis}
\end{tikzpicture}
\end{center}
\end{tcolorbox}

\begin{applicationexercise}{Classroom heat-loss budget}
A classroom has $120$ m$^2$ of wall with $U=0.25$ W\,m$^{-2}$\,K$^{-1}$ and $30$ m$^2$ of windows with
$U=1.4$ W\,m$^{-2}$\,K$^{-1}$. The indoor--outdoor temperature difference is $20$ K.
\begin{enumerate}
\item Calculate the transmission heat loss through walls and windows.
\item Add a ventilation loss of $0.65$ kW and a thermal-bridge allowance equal to 12\% of transmission loss.
\item A heat pump has COP $=3.5$. Estimate its electrical input power.
\item Propose one measured indicator that would reveal a commissioning problem after occupancy.
\end{enumerate}
\end{applicationexercise}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
embodied carbon & carbone incorporé\\
operational energy & énergie d’exploitation\\
airtightness & étanchéité à l’air\\
thermal bridge & pont thermique\\
shading & protection solaire\\
heat pump & pompe à chaleur\\
commissioning & mise au point\\
sub-metering & sous-comptage\\
performance gap & écart de performance\\
lifecycle assessment & analyse du cycle de vie
\end{tabularx}
\end{vocabulary}

\begin{questions}
\item What is the difference between operational and embodied carbon?
\item Why should demand reduction precede equipment selection?
\item How can external shading improve summer comfort?
\item Why can actual consumption differ from simulation?
\item Why is CO$_2$ useful but insufficient as a complete air-quality indicator?
\end{questions}

\begin{engineeringchallenge}
Design a low-energy classroom for a temperate climate. Set envelope, shading, ventilation and heating requirements;
calculate a peak heat-loss estimate; compare two structural options using embodied carbon; define five sensors and
three one-year performance indicators; and specify an acceptance test for comfort, air quality and energy use.
\end{engineeringchallenge}

\begin{speaking}
Should building regulations limit embodied carbon as strictly as operational energy? Discuss measurement uncertainty,
affordability, renovation and lifecycle responsibility.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{International Energy Agency -- Buildings}{https://www.iea.org/energy-system/buildings}
\ECsource{UN Environment Programme -- Buildings and construction}{https://www.unep.org/topics/cities/buildings-and-construction}